### 2020-10-31 Reports/JOURNAL - Confidence Regions/Chapters/Background.tex
subsection true random numbers

Let ${\mathcal X} \equiv \{x_t\}_{t=1}^{T}$ be a real valued time series of length $T$, without ties. 
As stated by \citeauthor{PermutationEntropyBandtPompe}~\ycite{PermutationEntropyBandtPompe} in their seminal work:  
\begin{quote}
``If the $\{x_t\}_{t=1}^{T}$ attain infinitely many values, it is common to replace them by a symbol sequence 
$\Pi \equiv \{\pi_j\}$ with finitely many symbols, and calculate source entropy from it".
\end{quote}
Also, as stressed by these authors, 
\begin{quote}
``The corresponding symbol sequence must come 
naturally from the $\{x_t\}_{t=1}^{T}$ without former model assumptions".
\end{quote}

Let ${\mathbbm A}_{D}$ (with $D \geq 2$ and $D \in {\mathbbm Z}$) be the symmetric group of order $D!$ formed by all 
possible permutation of order $D$, and the symbol component vector 
${\bm \pi}^{(D)} = (\pi_1, \pi_2, \dots, \pi_D)$ so every element ${\bm \pi}^{(D)}$ is unique 
($\pi_j \neq \pi_k~\forall~j \neq k$). 
Consider for the time series ${\mathcal X} \equiv \{x_t\}_{t=1}^{T}$ its time delay embedding representation,
with embedding dimension $D \geq 2$ ($D \in {\mathbbm Z}$) and time delay $\tau \geq 1$ ($\tau \in {\mathbbm Z}$, also called ``embedding time''):
\begin{equation} 
\label{eq:time-delay}
{\mathbf X}^{(D,\tau)}_t ~=~( x_t,x_{t+\tau},\dots,x_{t+(D-1)\tau} ) \ ,
\end{equation} 
for $t = 1,2,\dots,N$ with $N = T-(D-1) \tau$.
Then the vector ${\mathbf X}^{(D,\tau)}_t$ can be mapped to a symbol vector ${\bm \pi}^{(D)} \in {\mathbbm A}_{D}$. 
This mapping should be defined in a way that preserves the desired relation between the elements 
$x_t  \in {\mathbf X}^{(D,\tau)}_t$, and all $t \in T$ that share this pattern (also called motif) have to mapped to the same 
${\bm \pi}^{(D)}$. 
The two most frequent ways to define the mapping ${\mathbf X}^{(D,\tau)} \mapsto {\bm \pi}^{(D)}$ are:  
\begin{enumerate}[a)]
\item ordering the ranks of the $x_t \in {\mathbf X}^{(D,\tau)}$ in chronological order 
       (\textit{Rank Permutation}) or,
\item ordering the time indexes according to the ranks of $x_t \in {\mathbf X}^{(D,\tau)}$  
       (\textit{Chronological Index Permutation});
\end{enumerate}
       see details in \citeauthor{BPRepeatedValuesChaos}~\ycite{BPRepeatedValuesChaos}.
Without loss of generality, in the following we will only use the latter.

@@

Although very expressive, the Normalized Shannon Entropy is not able to describe all possible underlying dynamics.
In particular, for intermediate values of $H$, there is a wide variety of situations worth characterizing.
To this aim, \citeauthor{LopezRuiz1995}~\ycite{LopezRuiz1995} proposed using $Q$, the disequilibrium, a measure of how far $\bm h$ is from an equilibrium or noninformative distribution.
They employed the Euclidean distance between $\bm h$ and the uniform probability distribution function.

>>>>

Let ${\mathcal X} \equiv \{x_t\}_{t=1}^{T}$ be a real valued time series of length $T$, without ties. 
As stated by \Mycite{PermutationEntropyBandtPompe} in their seminal work:  
\begin{quote}
``If the $\{x_t\}_{t=1}^{T}$ attain infinitely many values, it is common to replace them by a symbol sequence 
$\Pi \equiv \{\pi_j\}$ with finitely many symbols, and calculate source entropy from it".
\end{quote}
Also, as stressed by these authors, 
\begin{quote}
``The corresponding symbol sequence must come 
naturally from the $\{x_t\}_{t=1}^{T}$ without former model assumptions".
\end{quote}

Let ${\mathbbm A}_{D}$ (with $D \geq 2$ and $D \in {\mathbbm Z}$) be the symmetric group of order $D!$ formed by all 
possible permutation of order $D$, and the symbol component vector 
${\bm \pi}^{(D)} = (\pi_1, \pi_2, \dots, \pi_D)$ so every element ${\bm \pi}^{(D)}$ is unique 
($\pi_j \neq \pi_k~\forall~j \neq k$). 
Consider for the time series ${\mathcal X} \equiv \{x_t\}_{t=1}^{T}$ its time delay embedding representation,
with embedding dimension $D \geq 2$ ($D \in {\mathbbm Z}$) and time delay $\tau \geq 1$ ($\tau \in {\mathbbm Z}$, also called ``embedding time''):
\begin{equation} 
\label{eq:time-delay}
{\mathbf X}^{(D,\tau)}_t ~=~( x_t,x_{t+\tau},\dots,x_{t+(D-1)\tau} ) \ ,
\end{equation} 
for $t = 1,2,\dots,N$ with $N = T-(D-1) \tau$.
Then the vector ${\mathbf X}^{(D,\tau)}_t$ can be mapped to a symbol vector ${\bm \pi}^{(D)} \in {\mathbbm A}_{D}$. 
This mapping should be defined in a way that preserves the desired relation between the elements 
$x_t  \in {\mathbf X}^{(D,\tau)}_t$, and all $t \in T$ that share this pattern (also called motif) have to mapped to the same 
${\bm \pi}^{(D)}$. 
The two most frequent ways to define the mapping ${\mathbf X}^{(D,\tau)} \mapsto {\bm \pi}^{(D)}$ are:  
\begin{enumerate}[label=\alph*)]
\item ordering the ranks of the $x_t \in {\mathbf X}^{(D,\tau)}$ in chronological order 
       (\textit{Rank Permutation}) or,
\item ordering the time indexes according to the ranks of $x_t \in {\mathbf X}^{(D,\tau)}$  
       (\textit{Chronological Index Permutation});
\end{enumerate}
       see details in \Mycite{BPRepeatedValuesChaos}.
Without loss of generality, in the following we will only use the latter.

@@

Although very expressive, the Normalized Shannon Entropy is not able to describe all possible underlying dynamics.
In particular, for intermediate values of $H$, there is a wide variety of situations worth characterizing.
To this aim, \Mycite{LopezRuiz1995} proposed using $Q$, the disequilibrium, a measure of how far $\bm h$ is from an equilibrium or noninformative distribution.
They employed the Euclidean distance between $\bm h$ and the uniform probability distribution function.

### 2020-10-31 Reports/JOURNAL - Confidence Regions/Chapters/ExperimentalSetup.tex
subsection true random numbers

\todo[inline]{Gerador - Marcelo}

Random numbers are used in many fields, from gambling to criptography, aiming to guarantee a secure, realistic or unpredictable behavihor. 
Pseudo randomic results can be achieved by software in a deterministic way. 
But, some applications need actual random numbers (despite the somewhat elusive nature of actual randomness).
Randomness can be observed in unpredictable real world phenomena like cathodic radiation or atmospheric noise.
In this study we used two sources of real random numbers. 
The first is based on vacuum states to generate random quantum numbers described by \citeauthor{RNGVacuumStates}~(\citeyear{RNGVacuumStates}), the second one is based on atmospheric noise captured by a cheap radio receiver presented at \url{www.random.org}.

>>>>

Random numbers are used in many fields, from gambling to criptography, aiming to guarantee a secure, realistic or unpredictable behavior. 
Pseudo randomic results can be achieved by software in a deterministic way. 
But, some applications need actual random numbers (despite the somewhat elusive nature of actual randomness).
Randomness can be observed in unpredictable real world phenomena like cathodic radiation or atmospheric noise.
%In this study we used two sources of real random numbers. 
%The first is based on vacuum states to generate random quantum numbers described by \citeauthor{RNGVacuumStates}~(\citeyear{RNGVacuumStates}), the second one is based on atmospheric noise captured by a cheap radio receiver presented at \url{www.random.org}.

In this study we used two sources of random numbers, here called true random, both from physical phenomena observation and measurement.
The first is based on vacuum states to generate random quantum numbers, the setup consists of an ordinary laser source to generate a local oscillator (LO), a half-wave plate, a polarizing beamsplitter (BPS), and two balanced detectors working together adding or subtracting the photocurrents results in a quadrature measurement of the LO or vacuum state. 
The probability distribution of the vacuum state is binned into $2^n$ equal parts (bins of same size), than, assigning a fixed bit combination of length \textit{n} to each sample point in a given bin \citeauthor{RNGVacuumStates}~(\citeyear{RNGVacuumStates}). 
The second one is based on atmospheric noise captured by a cheap radio receiver, started as a gambling engine, the randomness comes from an ordinary radio receiver that has no filter for static unwanted sounds caused by atmospheric noise, but perfect for random purposes, developed over a distributed setup with some radios located at different geographical locations sending random bits to a cloud server who process data and hosts random.org, the history, and some other information could be found at \cite{RandomOrg}.

### 2020-10-31 Reports/JOURNAL - Confidence Regions/Chapters/Introduction.tex
subsection true random numbers

Time Series carry valuable information about the system which produces the data.
Their analysis is usually based on two approaches \cite{TimeSeriesAnalysisCryerChan}: in the (natural) time and transformed domains (for instance, frequency and wavelet).
In the context of time-domain analysis, a new methodology was proposed by \citeauthor{PermutationEntropyBandtPompe}~\ycite{PermutationEntropyBandtPompe}.
Its approach is non-parametric and based on descriptors of Information Theory.
Through this, the time series is transformed into ordinal patterns, with which a histogram is formed.
Using such patterns, the resulting distribution becomes less sensitive to outliers and, as it does not depend on any model, can be applied to a variety of situations.

@@

Each time series is described by a point in the range of $\mathbbm R^2$, the entropy complexity plane.
Two points are well known in this plane: those of white noise and a completely deterministic sequence.
Through these references, we can characterize the time series according to the dynamics of its generating process.
Based on this premise, studies with different applications managed to obtain relevant results from time series through information on the nature of the data provided by the $H \times C$ plane.
Examples include the analysis of \citeauthor{echegoyen2020permutation}~\ycite{echegoyen2020permutation} in magnetoencephalography recordings of individuals suffering from mild cognitive impairment and individuals diagnosed with Alzheimer's disease by trajectories in the $H \times C$ plane,
\citeauthor{InformationTheoryPerspectiveNetworkRobustness}~\ycite{InformationTheoryPerspectiveNetworkRobustness} verified the effect of attacks on complex networks by displacing their points in the $H \times C$ plane,
\citeauthor{CharacterizationVehicleBehaviorInformationTheory}~\ycite{CharacterizationVehicleBehaviorInformationTheory} described the behavior of vehicles depending on the topology of cities, and
\citeauthor{Chagas2020Characterization}~\ycite{Chagas2020Characterization} succeeded in expanding the use of such techniques for analyzing SAR texture images, making characterization, and classification of them.

@@

In the context of theoretical work, chaotic and random component analysis has been developed with the aim of improving the understanding of the plane's properties. We can cite as highlights:
\citeauthor{GeneralizedStatisticalComplexityMeasuresGeometricalAnalyticalProperties}~\ycite{GeneralizedStatisticalComplexityMeasuresGeometricalAnalyticalProperties} analyzed the logistic chaotic map and discuss the boundaries of the $H \times C$ plane.
\citeauthor{De_Micco_2009}~\ycite{De_Micco_2009} studied chaotic components in pseudo-random number generators.
\citeauthor{DistinguishingNoiseFromChaos}~\ycite{DistinguishingNoiseFromChaos}  tackled the often hard problem of distinguishing chaos and noise.
\citeauthor{DistinguishingChaoticStochasticDynamicsTimeSeriesMultiscaleSymbolicApproach}~\ycite{DistinguishingChaoticStochasticDynamicsTimeSeriesMultiscaleSymbolicApproach} used a multi-scale approach to analyze the interplay between chaotic and stochastic dynamics.
With the knowledge of the expected variability of such points, according to the underlying dynamics, we can make hypothesis tests for a wide variety of models.
Some preliminary results can already be found in the literature.
\citeauthor{RandomNumberGeneratorsCausality}~\ycite{RandomNumberGeneratorsCausality} showed that the Entropy-Complexity plane ($H\times C$) is a good indicator of the results of Diehard tests for pseudorandom number generators.
\citeauthor{De_Micco_2008}~\ycite{De_Micco_2008} assessed ways of improving pseudorandom sequences by their representation in this plane.
\citeauthor{LiborInvisibleHand}~\ycite{LiborInvisibleHand} identified spurious interventions in the Libor market using the $H\times C$ plane representation.

>>>>

Time Series carry valuable information about the system which produces the data.
Their analysis is usually based on two approaches \cite{TimeSeriesAnalysisCryerChan}: in the (natural) time and transformed domains (for instance, frequency and wavelet).
In the context of time-domain analysis, a new methodology was proposed by \Mycite{PermutationEntropyBandtPompe}.
Its approach is non-parametric and based on descriptors of Information Theory.
Through this, the time series is transformed into ordinal patterns, with which a histogram is formed.
Using such patterns, the resulting distribution becomes less sensitive to outliers and, as it does not depend on any model, can be applied to a variety of situations.

@@

Each time series is described by a point in the range of $\mathbbm R^2$, the entropy complexity plane.
Two points are well known in this plane: those of white noise and a completely deterministic sequence.
Through these references, we can characterize the time series according to the dynamics of its generating process.
Based on this premise, studies with different applications managed to obtain relevant results from time series through information on the nature of the data provided by the $H \times C$ plane.
Examples include the analysis of \Mycite{echegoyen2020permutation} in magnetoencephalography recordings of individuals suffering from mild cognitive impairment and individuals diagnosed with Alzheimer's disease by trajectories in the $H \times C$ plane,
\Mycite{InformationTheoryPerspectiveNetworkRobustness} verified the effect of attacks on complex networks by displacing their points in the $H \times C$ plane,
\Mycite{CharacterizationVehicleBehaviorInformationTheory} described the behavior of vehicles depending on the topology of cities, and
\Mycite{Chagas2020Characterization} succeeded in expanding the use of such techniques for analyzing SAR texture images, making characterization, and classification of them.

@@

In the context of theoretical work, chaotic and random component analysis has been developed with the aim of improving the understanding of the plane's properties. We can cite as highlights:
\Mycite{GeneralizedStatisticalComplexityMeasuresGeometricalAnalyticalProperties} analyzed the logistic chaotic map and discuss the boundaries of the $H \times C$ plane.~\citeyear{De_Micco_2009} studied chaotic components in pseudo-random number generators.
\Mycite{DistinguishingNoiseFromChaos}  tackled the often hard problem of distinguishing chaos and noise.
\Mycite{DistinguishingChaoticStochasticDynamicsTimeSeriesMultiscaleSymbolicApproach} used a multi-scale approach to analyze the interplay between chaotic and stochastic dynamics.
With the knowledge of the expected variability of such points, according to the underlying dynamics, we can make hypothesis tests for a wide variety of models.
Some preliminary results can already be found in the literature.
\Mycite{RandomNumberGeneratorsCausality} showed that the Entropy-Complexity plane ($H\times C$) is a good indicator of the results of Diehard tests for pseudorandom number generators.
\Mycite{De_Micco_2008} assessed ways of improving pseudorandom sequences by their representation in this plane.
\Mycite{LiborInvisibleHand} identified spurious interventions in the Libor market using the $H\times C$ plane representation.

### 2020-10-31 Reports/JOURNAL - Confidence Regions/Chapters/relatedWork.tex
subsection true random numbers

The first works on the characterization of white noises with permutation entropy arose from the need to discriminate them in relation to chaotic maps~\cites{rosso2013characterization, xiong2020complexity, olivares2012contrasting}.
However, it was found that the measure of statistical complexity was able to efficiently quantify the performance of pseudorandom number generators, expanding the possibilities of using information theory descriptors with the Bandt-Pompe symbolization~\cites{larrondo2002statistical, gonzalez2005statistical}.

>>>>

The first works on the characterization of white noises with permutation entropy arose from the need to discriminate them in relation to chaotic maps~\cite{rosso2013characterization, xiong2020complexity, olivares2012contrasting}.
However, it was found that the measure of statistical complexity was able to efficiently quantify the performance of pseudorandom number generators, expanding the possibilities of using information theory descriptors with the Bandt-Pompe symbolization~\cite{larrondo2002statistical, gonzalez2005statistical}.
